Large Language Models in Reinforcement Learning - A Comparison of Methods

### 2023-11-14 doc/paper.tex
Update paper
Signed-off-by: Jonatan Hanssen <jonatan.hanssen@outlook.com>

Reinforcement Learning (RL) algorithms suffer when rewards are sparse and the state-action space is large. Even tasks which seem relatively simple can prove intractable if the completion of the task requires subtasks to be completed first, or if the reward only comes when the entire task has been completed. In such cases, random exploration is unlikely to lead the agent to discover a solution to the problem. The apparent simplicity of such problems is often due to human intuition, which allow us to quickly see possible solutions to a diverse set of problems. Large Language Models (LLMs) are trained on large corpora of human written text, and have been shown to capture parts of this intuition in many tasks. In recent years, LLMs have been successfully used to aid RL agents in more efficient exploration, and have allowed them to solve problems which previous methods have been unable to. We compare different methods for integrating an LLM into the Deep Learning RL algorithm Proximal Policy Optimization (PPO), and compare their efficiency against each other. \footnote{Code available at \href{https://github.com/jonatan-hanssen/in5490-project}{GitHub}}

@@

RL informed by natural language is a relatively new field, where research before 2018 have been limited to relatively small corpora or synthetic language \cite[1]{survey}. With the introduction of Large Language Models, many new possibilities have opened up, and in recent years there have been many successful integrations of these models into traditional RL \cite{omni, ellm, idm}.

@@

The algorithm proposed in this paper involves prompting the LLM with the agent's state observation, a list of tasks the agent already does well, and a list of available tasks for the LLM to deem as either interesting or boring. After sorting available tasks into interesting and boring ones, it assigns sampling weights to them, giving much higher weights to interesting tasks than boring ones, thus directly influencing the agent's policy by altering probabilities of which action it then chooses to do. In \cite{idm}, another paper proposing using the LLM as a policy influencer, has a similar approach in which the state observation is tokenized and passed to the LLM, though in their experiment, goals and action histories are also tokenized and passed. Here, too, the LLM response is fed back into a task-specific RL model, which in turn influences policy probabilities of certain actions being taken. In general, many of the novel methods suggested involve the LLM suggesting actions for the agent to perform in a given context, and its output influencing the agent policy.

>>>>

Reinforcement Learning (RL) algorithms suffer when rewards are sparse and the state-action space is large. Even tasks which seem relatively simple can prove intractable if the completion of the task requires subtasks to be completed first, or if the reward only comes when the entire task has been completed. In such cases, random exploration is unlikely to lead the agent to discover a solution to the problem. The apparent simplicity of such problems is often due to human intuition, which allow us to quickly see possible solutions to a diverse set of problems. Large Language Models (LLMs) are trained on large corpora of human written text, and have been shown to capture parts of this intuition in many tasks. In recent years, LLMs have been successfully used to aid RL agents in more efficient exploration, and have allowed them to solve problems which previous methods have been unable to. We compare different methods for integrating an LLM into the Deep Learning RL algorithm Proximal Policy Optimization (PPO), and compare their efficiency against each other.\footnote{Code available at \href{https://github.com/jonatan-hanssen/in5490-project}{GitHub}}

@@

RL informed by natural language is a relatively new field, where research before 2018 had been limited to relatively small corpora or synthetic language \cite[1]{survey}. With the introduction of Large Language Models, many new possibilities have opened up, and in recent years there have been many successful integrations of these models into traditional RL \cite{omni, ellm, idm}.

@@

The algorithm proposed in this paper involves prompting the LLM with the agent's state observation, a list of tasks the agent already does well, and a list of available tasks for the LLM to deem as either "interesting" or "boring". After sorting available tasks into interesting and boring ones, it assigns sampling weights to them, giving much higher weights to interesting tasks than boring ones. Thus, it directly influences the agent's policy by altering the probabilities of the possible actions. In \cite{idm}, another paper proposing using the LLM as a policy influencer, has a similar approach in which the state observation is tokenized and passed to the LLM, though in their experiment, goals and action histories are also tokenized and passed. Here, too, the LLM response is fed back into a task-specific RL model, which in turn influences policy probabilities of certain actions being taken. In general, many of the novel methods suggested involve the LLM suggesting actions for the agent to perform in a given context, and its output influencing the agent policy.
